CrypTool Script \\ Mathematics and Cryptography

### 2003-03-10 script/en/chapters/aboutcryptool.tex
no message

You will receive information about the principles of symmetrical and
asymmetrical encryption. A large section of the script is dedicated to the
fascinating topic of prime numbers. Using numerous examples, the elementary number
theory and modular arithmetic are introduced and applied in an exemplary 
manner in the RSA procedure.  After this, one receives insight into the mathematical ideas
behind modern cryptography.

@@

A further chapter is devoted to digital signatures, which are an essential
component of e-business applications. The last chapter --- elliptic curves ---
fits in well with this. The mathematics of elliptic curves forms the basis for
rapid cryptographic algorithms for digital signatures; these algorithms are
extremely well suited for implementation on smartcards.

@@

The {\em authors} Bernhard Esslinger, Mathias B"uger,Bartol Filipovic, 
Henrik Koy, Roger Oyono and J\"org Cornelius Schneider
would like to take this opportunity to thank their colleagues in the company and
at the universities of Frankfurt, Gie\ss en, Siegen and Karlsruhe. They particularly thank
Dr. Peer Wichmann from the Karlsruhe computer science research centre
(Forschungszentrum Informatik, FZI) for his down-to-earth support.

\enlargethispage{0.5cm}
As at CrypTool, the quality of the script is enhanced through your suggestions
and ideas for improvement. We look forward to your feedback.

\vskip +7 pt \noindent
You will find the current version of CrypTool\index{CrypTool} under \newline
  \href{http://www.CrypTool.de}{\texttt{http://www.cryptool.de}},~
  \href{http://www.cryptool.com}{\texttt{http://www.cryptool.com}}~ or~
  \href{http://www.cryptool.org}{\texttt{http://www.cryptool.org}}.
{
\vskip + 7 pt \noindent
The contact persons for this free tool are listed in the readme file for
CrypTool.}
}

>>>>

You will receive information about the principles of symmetrical and
asymmetrical {\em encryption}. A large section of the script is dedicated
to the fascinating topic of {\em prime numbers}. Using numerous examples,
the {\em elementary number theory} and {\em modular arithmetic} are
introduced and applied in an exemplary manner to the RSA procedure.
After this, one receives insight into the mathematical ideas behind
{\em modern cryptography}.

@@

A further chapter is devoted to {\em digital signatures}, 
which are an essential component of e-business applications. 
The last chapter describes {\em elliptic curves}: they could be used
as an alternative to RSA they are extremely well suited for 
implementation on smartcards.

@@

The {\em authors} Bernhard Esslinger, Matthias B"uger, Bartol Filipovic, 
Henrik Koy, Roger Oyono and J\"org-Cornelius Schneider
would like to take this opportunity to thank their colleagues 
in the company and at the universities of Frankfurt, Gie\ss en, 
Siegen and Karlsruhe. They particularly thank Dr. Peer Wichmann
from the Karlsruhe computer science research centre
(Forschungszentrum Informatik, FZI) for his down-to-earth support.

\enlargethispage{0.5cm}
As at CrypTool\index{CrypTool}, the quality of the script is enhanced
through your suggestions and ideas for improvement. We look forward 
to your feedback.

\vskip +7 pt \noindent
You will find the current version of CrypTool\index{CrypTool} under \newline
  \href{http://www.cryptool.org}{\texttt{http://www.cryptool.org}},~
  \href{http://www.cryptool.com}{\texttt{http://www.cryptool.com}}~ or~
  \href{http://www.CrypTool.de}{\texttt{http://www.cryptool.de}}.
{
\vskip + 7 pt \noindent
The contact persons for this free open source tool are listed 
in the readme file delivered within the CrypTool\index{CrypTool} package.}
}

### 2003-03-10 script/en/chapters/cryptomethods.tex
no message

% ..............................................................................
%                V E R S C H L U E S S E L U N G S V E R F A H R E N
% ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

\newpage
\section{Encryption procedures}
\hypertarget{Kapitel_1}{}
(Bernhard Esslinger, besslinger@web.de, May 1999, Updates Dec. 2001, Feb. 2003)

@@

% --------------------------------------------------------------------------
\subsubsection[Symmetric encryption]
{Symmetric encryption\footnotemark}
\footnotetext{%
With CrypTool\index{CrypTool} v1.3 you can execute the following modern
symmetric encryption algorithms 
(using the menu ``Crypt \textbackslash Symmetric''): \\
IDEA, RC2, RC4, DES (ECB), DES (CBC), Triple-DES (ECB), Triple-DES (CBC),
MARS (AES candidate), RC6 (AES candidate), Serpent (AES candidate), 
Twofish (AES candidate), Rijndael (official AES algorithm).
}

@@

More details about the AES algorithms can be found within the Online help of CrypTool\index{CrypTool} (in the index see head-word {\em AES} and then the help pages {\em AES candidates}, {\em The AES Winner Rijndael} and {\em The Rijndael encryption algorithm}).

@@

Below you will find some results, which have recently called into question the security of the AES algorithm -- from our point of view these doubts practically still remain unfounded. The following information is based on the original papers and the articles \cite{Wobst-iX2002} and \cite{Lucks-DuD2002}.

@@

More details can be found at: \\
\href{http://www.cryptosystem.net/aes}{\texttt{http://www.cryptosystem.net/aes}} and \\ 
\href{http://www.minrank.org/aes/}{\texttt{http://www.minrank.org/aes/}}

So for 256-AES the attack is much more effective than brute-force but still far away from any computing power which could be accessible in the short-to-long term. The discussion is very controversial: Don Coppersmith (one of the inventors of DES) for example queries the practicability of the attack because XLS would provide no solution for AES \cite{Coppersmith2002}. This concludes that then the optimization of Murphy and Robshaw \cite{Robshaw2002b} would not work.

% --------------------------------------------------------------------------
\subsubsection{Asymmetric encryption}

@@

% --------------------------------------------------------------------------
\newpage
\subsubsection[Hybrid procedures]
{Hybrid procedures\footnotemark}
\footnotetext{%
Within CrypTool\index{CrypTool} v1.3 you can get a visualization of this
technique using the menu ``Crypt \textbackslash Hybrid Demonstration'': 
this dialogue shows the single steps and its dependencies with concrete
numbers.
}\index{Hybrid procedures}

@@

In this case the data is encrypted using symmetric procedures: the key is a
session key\index{Session key} generated by the sender randomly that is only used for this message.
This session key is then encrypted using the asymmetric procedure and
transmitted to the recipient together with the message. Recipients can determine
the session key using their secret keys and then use the session key to encrypt
the message. In this way, we can benefit from the easy key management
\index{Key management} of asymmetric procedures and encrypt large quantities of data
quickly and efficiently using symmetric procedures.

@@

% --------------------------------------------------------------------------
\begin{thebibliography}{99999}
\addcontentsline{toc}{subsection}{Bibliography}

\bibitem[Schmeh2003]{Schmeh2003}  \index{Schmeh 2003}
        Klaus Schmeh, \\
        {\em Cryptography and Public Key Infrastructures on the Internet}, 
	John Wiley \& Sons Ltd., Chichester 2003. \\
        A considerable, up-to-date, good reading book, which also 
	considers practical problems, like standardisation or
        real existing software.

\bibitem[Coppersmith2002]{Coppersmith2002}  \index{Coppersmith 2002}
        Don Coppersmith, \\
        {\em Re: Impact of Courtois and Pieprzyk results}, 
	2002-09-19, AES Discussion Groups at \\
        \href{http://aes.nist.gov/aes/}
        {\texttt{http://aes.nist.gov/aes/}}

\bibitem[Courtois2002]{Courtois2002}  \index{Courtois 2002}
        Nicolas Courtois, Josef Pieprzyk, \\
        {\em Cryptanalysis of Block Ciphers with Overdefined Systems of Equations}, 
	received 10 Apr 2002, last revised 9 Nov 2002,
	A different version, so called compact version of the first XSL attack,
	was published in Asiacrypt Dec 2002, \\
        \href{http://eprint.iacr.org/2002/044}
        {\texttt{http://eprint.iacr.org/2002/044}}

@@

>>>>

% ..............................................................................
%                V E R S C H L U E S S E L U N G S V E R F A H R E N
%
% Schreibregelung: Capitel header to capitalize words,
%                  All other headers in normal lower/upper case 
%                  like sentences without dot at the end.
%
% ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

\newpage
\section{Encryption Procedures}
\hypertarget{Kapitel_1}{}
(Bernhard Esslinger, besslinger@web.de, May 1999, Updates Dec. 2001, Feb. 2003)

@@

% --------------------------------------------------------------------------
\subsubsection[Symmetric encryption]
{Symmetric encryption\footnotemark}
  \footnotetext{%
    With CrypTool\index{CrypTool} v1.3 you can execute the following modern
    symmetric encryption algorithms 
    (using the menu {\bf Crypt \textbackslash{} Symmetric}): \\
    IDEA, RC2, RC4, DES (ECB), DES (CBC), Triple-DES (ECB), Triple-DES (CBC),
    MARS (AES candidate), RC6 (AES candidate), Serpent (AES candidate), 
    Twofish (AES candidate), Rijndael (official AES algorithm).
  }

@@

More details about the AES algorithms can be found within the 
Online help of CrypTool\index{CrypTool}
  \footnote{%
      CrypTool Online help\index{CrypTool}: the index see head-word {\bf AES}
      leads to the help pages: {\bf AES candidates}, 
      {\bf The AES Winner Rijndael} and 
      {\bf The Rijndael encryption algorithm}.
  }
.

@@

Below you will find some results, which have recently called into question the security of the AES algorithm -- from our point of view these doubts practically still remain unfounded
% = do not bring disrepute upon AES
. 
The following information is based on the original papers and the articles \cite{Wobst-iX2002} and \cite{Lucks-DuD2002}.

@@

More details can be found at: \\
\begin{itemize}
  \item[] \href{http://www.cryptosystem.net/aes}
               {\texttt{http://www.cryptosystem.net/aes}}
  \item[] \href{http://www.minrank.org/aes/}
               {\texttt{http://www.minrank.org/aes/}}
\end{itemize}

So for 256-AES the attack is much more effective than brute-force but still far more away from any computing power which could be accessible in the short-to-long term.

The discussion is very controversial: Don Coppersmith (one of the inventors of DES) for example queries the practicability of the attack because XLS would provide no solution for AES \cite{Coppersmith2002}. This concludes that then the optimization of Murphy and Robshaw \cite{Robshaw2002b} would not work.

% --------------------------------------------------------------------------
\subsubsection[Asymmetric encryption]
{Asymmetric encryption\footnotemark}
  \footnotetext{%
    With CrypTool\index{CrypTool} v1.3 you can execute RSA encryption
    and decryption (using the menu {\bf Crypt \textbackslash{} Asymmetric}).
  }

@@

% --------------------------------------------------------------------------
% \newpage
\subsubsection[Hybrid procedures]
{Hybrid procedures\footnotemark}
\footnotetext{%
Within CrypTool\index{CrypTool} v1.3 you can get a visualization of this
technique using the menu {\bf Crypt \textbackslash{} Hybrid Demonstration}: 
this dialogue shows the single steps and its dependencies with concrete
numbers.
}\index{Hybrid procedures}

@@

In this case the data is encrypted using symmetric procedures: the key is a
session key\index{Session key} generated by the sender
randomly \footnote{%
An important part of cryptographically secure techniques is to generate 
random numbers. Within CrypTool\index{CrypTool} v1.3 you can check out
different random number generators using the menu {\bf Indiv. Procedures 
\textbackslash{} Generate Random Numbers}. 
Using the menu{\bf Analysis \textbackslash{} Random Tests} you can apply
different test methods for random data to binary documents. \\
Till now CrypTool concentrates on cryptographically strong 
pseudo number generators. Only the integrated Secude generator 
involves a "pure" random source. 
}\index{Random}
that is only used for this message.
This session key is then encrypted using the asymmetric procedure and
transmitted to the recipient together with the message. Recipients can determine
the session key using their secret keys and then use the session key to encrypt
the message. In this way, we can benefit from the easy key management
\index{Key management} of asymmetric procedures and encrypt large quantities of data
quickly and efficiently using symmetric procedures.

@@

% --------------------------------------------------------------------------
\newpage
\begin{thebibliography}{99999}
\addcontentsline{toc}{subsection}{Bibliography}

\bibitem[Schmeh2003]{Schmeh2003}  \index{Schmeh 2003}
        Klaus Schmeh, \\
        {\em Cryptography and Public Key Infrastructures on the Internet},\\ 
	John Wiley \& Sons Ltd., Chichester 2003. \\
        A considerable, up-to-date, good reading book, which also 
	considers practical problems like standardisation or
        real existing software.

\bibitem[Coppersmith2002]{Coppersmith2002}  \index{Coppersmith 2002}
        Don Coppersmith, \\
        {\em Re: Impact of Courtois and Pieprzyk results}, \\
	2002-09-19, ``AES Discussion Groups''~ at \\
        \href{http://aes.nist.gov/aes/}
        {\texttt{http://aes.nist.gov/aes/}}

\bibitem[Courtois2002]{Courtois2002}  \index{Courtois 2002}
        Nicolas Courtois, Josef Pieprzyk, \\
        {\em Cryptanalysis of Block Ciphers with Overdefined Systems of Equations}, \\
	received 10 Apr 2002, last revised 9 Nov 2002.\\
	A different version, so called compact version of the first XSL attack,
	was published at Asiacrypt Dec 2002. \\
        \href{http://eprint.iacr.org/2002/044}
        {\texttt{http://eprint.iacr.org/2002/044}}

@@

% --------------------------------------------------------------------------
\newpage
\section*{Web links}\addcontentsline{toc}{subsection}{Web links}

\begin{enumerate}
  \item[] \href{http://www.cryptosystem.net/aes}
               {\texttt{http://www.cryptosystem.net/aes}}
  \item[] \href{http://www.minrank.org/aes/}
               {\texttt{http://www.minrank.org/aes/}}
\end{enumerate}

### 2003-03-10 script/en/chapters/digitalsignatures.tex
no message

\newpage
\section{Digital signatures}
\index{Signatures!digital}
(Esslinger B. / Filipovics B. / Schneider J. / Koy H., Update: June 2002)

@@

An asymmetric technique is used again (see encryption procedures). Participants
who wish to generate a digital signature for a document must possess a pair of
keys. They use their secret key to generate signatures and the recipient uses
the sender's public key to verify whether the signature is correct. As before,
it must be impossible to use the public key to derive the secret key\footnote{%
With CrypTool\index{CrypTool} v1.3 you can also generate and check 
digital signatures: \\
using the submenus of the main menu ``Digital Signatures'' or \\
using menu {\bf Indiv. Procedures \textbackslash RSA Demonstration 
\textbackslash Signature Demonstration}.
}.

@@

The document is sent to the recipient together with the signature. The recipient
can then use the sender's public key, the document and the signature to
establish whether or not the signature is correct. In practice, however, the
procedure we have just described has a decisive disadvantage. The signature is
approximately as long as the document itself. To prevent an unnecessary increase
in data traffic, and also for reasons of performance\index{Performance}, we use 
a cryptographic hash function\footnote{%
%xxxxxxxxxx Menüeinträgsbezeichnungen ok?
Hash functions are implemented within CrypTool\index{CrypTool} v1.3.04 at
several places: \\
using menu  ``Indiv. Procedures \textbackslash Hash'' you have the possibility
\begin{itemize}\addtolength{\itemsep}{-1.0\baselineskip} % um den Abstand zu verringern bebe
\item to apply one of 6 hash functions to the content of the current window, \\
\item to calculate the hash value of a file, \\
\item to test, how changes to a text change the according hah value and \\
\item to perform a simulation, how digital signatures could be attacked 
      by a targeted search for hash value collisions.
\end{itemize}
}
.

@@

\subsection{RSA Signatures}
\index{Signatures!digital}
\index{RSA signatures}

@@

\subsection{DSA Signatures}
\index{Signatures!digital}
\index{DSA signatures}

@@

\paragraph{Signing}\strut\\
\begin{tabular}{l@{ }l}
$m$ & the message to be signed\\
$k$ & choose at random, less than $q$\\
$r$ & $= (g^k \mbox{ mod } p) \mbox{ mod } q$\\
$s$ & $= (k^{-1}(\mbox{SHA-1}(m) + xr)) \mbox{ mod } q$
\end{tabular}

\emph{Remark:}
\begin{itemize}
\item $(s,r)$ is the signature.
\item The security of the signature depends not only on the mathematical
  properties, but also on using a good random source  for $k$.
\item SHA-1 \index{SHA-1} is a 160-bit hash function specified in FIPS186.
\end{itemize}
\paragraph{Verifying}\strut\\
\begin{tabular}{l@{ }l}
$w$ & $= s^{-1}  \mbox{ mod } q$\\
$u_1$ & $= (\mbox{SHA-1}(m)w) \mbox{ mod } q$\\
$u_2$ & $= (rw)  \mbox{ mod } q$\\
$v$ & $= (g^{u_1}y^{u_2}) \mbox{ mod } p)  \mbox{ mod } q$\\

@@

\begin{thebibliography}{99999}
\addcontentsline{toc}{subsection}{Bibliography}
\bibitem[Schneier1996]{5Schneier1996} \index{Schneier 1996} 
    Bruce Schneier, \\
    Applied Cryptography, Protocols, Algorithms,
    and Source Code in C, Wiley, 2nd edition, 1996.

        \bibitem[PKCS1]{PKCS1} RSA Laboratories\\ \index{PKCS\#1}
          PKCS \#1 v2.1 Draft 3: RSA Cryptography Standard. April 19, 2002.

        \bibitem[FIPS186]{FIPS186} U.S. Department of
          Commerce/N.I.S.T.\\\index{FIPS186}
          Entity authentication using public key cryptography. February 18, 1997.
\end{thebibliography}

>>>>

\newpage
\section{Digital Signatures}
\index{Signatures!digital}
(Esslinger B. / Filipovics B. / Schneider J. / Koy H., June 2002,
Update: Feb. 2003)

@@

An asymmetric technique is used again (see encryption procedures). Participants
who wish to generate a digital signature for a document must possess a pair of
keys. They use their secret key to generate signatures and the recipient uses
the sender's public key to verify whether the signature is correct. As before,
it must be impossible to use the public key to derive the secret key\footnote{%
With CrypTool\index{CrypTool} v1.3 you can also generate and check 
digital signatures: \\
using the submenus of the main menu {\bf Digital Signatures} or \\
using menu {\bf Indiv. Procedures \textbackslash{} RSA Demonstration 
\textbackslash{} Signature Demonstration}.
}.

@@

The document is sent to the recipient together with the signature. The recipient
can then use the sender's public key, the document and the signature to
establish whether or not the signature is correct. In practice, however, the
procedure we have just described has a decisive disadvantage. The signature is
approximately as long as the document itself. To prevent an unnecessary increase
in data traffic, and also for reasons of performance\index{Performance}, we use 
a cryptographic hash function\footnote{%
%xxxxxxxxxx Menüeinträgsbezeichnungen ok?
Hash functions\index{Hash function} are implemented within CrypTool\index{CrypTool} v1.3.04 at
several places. \\
Using menu {\bf Indiv. Procedures \textbackslash{} Hash} you have the possibilities
\begin{itemize}\addtolength{\itemsep}{-1.0\baselineskip} % um den Abstand zu verringern bebe
\item to apply one of 6 hash functions to the content of the current window, \\
\item to calculate the hash value of a file, \\
\item to test, how changes to a text change the according hah value and \\
\item to perform a simulation, how digital signatures could be attacked 
      by a targeted search for hash value collisions.
\end{itemize}
}
.

@@

\subsection{RSA signatures}
\index{Signatures!digital}
\index{RSA signatures}

@@

\subsection{DSA signatures}
\index{Signatures!digital}
\index{DSA signatures}

@@

\paragraph{Signing}\strut\\
\begin{tabular}{l@{ }l}
$m$ & the message to be signed\\
$k$ & choose at random\index{Random}, less than $q$\\
$r$ & $= (g^k \mbox{ mod } p) \mbox{ mod } q$\\
$s$ & $= (k^{-1}(\mbox{SHA-1}(m) + xr)) \mbox{ mod } q$
\end{tabular}

\emph{Remark:}
\begin{itemize}
\item $(s,r)$ is the signature.
\item The security of the signature depends not only on the mathematical
  properties, but also on using a good random source\index{Random}  for $k$.
\item SHA-1 \index{SHA-1} is a 160-bit hash function specified in FIPS186.
\end{itemize}
\paragraph{Verifying}\strut\\
\begin{tabular}{l@{ }l}
$w$ & $= s^{-1}  \mbox{ mod } q$\\
$u_1$ & $= (\mbox{SHA-1}(m)w) \mbox{ mod } q$\\
$u_2$ & $= (rw)  \mbox{ mod } q$\\
$v$ & $= (g^{u_1}y^{u_2}) \mbox{ mod } p)  \mbox{ mod } q$\\

@@

\begin{thebibliography}{99999}
\addcontentsline{toc}{subsection}{Bibliography}

\bibitem[Schneier1996]{5Schneier1996} \index{Schneier 1996} 
    Bruce Schneier, \\
    Applied Cryptography, Protocols, Algorithms, and Source Code in C, \\
    Wiley, 2nd edition, 1996.

\bibitem[PKCS1]{PKCS1} RSA Laboratories\\ \index{PKCS\#1}
    PKCS \#1 v2.1 Draft 3: RSA Cryptography Standard, April 19, 2002.

\bibitem[FIPS186]{FIPS186} U.S. Department of Commerce/N.I.S.T.
    \index{FIPS186} \\
    Entity authentication using public key cryptography, February 18, 1997.

\end{thebibliography}

### 2003-03-10 script/en/chapters/ellipticcurves.tex
no message

\newpage
\section{Elliptic Curves}
\index{Elliptic curves}
\hypertarget{ellcurve}{}
(Filipovics B. / Esslinger B. / Oyono R., April 2000, Updates: Dec. 2001,
June 2002)

@@

The secret key $s$ of the signature generator is a (random) whole number $s$ in
the interval $[1, r-1]$. The public key of the signature generator is a point
$W=(x,y)$ on the elliptic curve ${\bf E}$. The public key $W$ and secret key $s$
are interrelated as follows: $W = sG$. This means that the domain parameters
(particularly $G$) and the secret key $s$ are used to calculate the public key
$W$ (by adding $G$ $s$ times on ${\bf E}$). The ECDLP is obviously used here: If
$W$ and $G$ (as well as the other domain parameters used) are known, then it is
difficult to use these to calculate $s$. (If the parameters are chosen
correctly, this currently appears to be practically impossible).

@@

CrypTool\index{CrypTool} uses elliptic curves for the digital signature function.

@@

\newpage
%\addcontentsline{toc}{subsection}{Literaturverzeichnis}
\begin{thebibliography}{99999}
\addcontentsline{toc}{subsection}{Bibliography}
        \bibitem[Lenstra1987]{Lenstra1987} H.W. Lenstra\\ \index{Lenstra 1987}
                Factoring integers with elliptic curves, Annals of Mathematics 126, pp. 649-673, 1987.
\end{thebibliography}
%\newpage

\section*{Web links}\addcontentsline{toc}{subsection}{Web links}

\begin{enumerate}
   \item Certicom Online Tutorial, \index{Certicom}\\
                \href{http://www.certicom.com/resources/ecc_tutorial/ecc_tutorial.html}{\texttt{http://www.certicom.com/resources/ecc\_tutorial/ecc\_tutorial.html}}
        \item IEEE P1363

                \href{http://grouper.ieee.org/groups/1363}{\texttt{http://grouper.ieee.org/groups/1363}}
        \item  \hypertarget{Lenstra2}{}
An informative web page about factorisation with elliptic curves.

\href{http://www.loria.fr/~zimmerma/records/ecmnet.html}{\texttt{http://www.loria.fr/\~{}zimmerma/records/ecmnet.html}}

There, one finds literature to the topic factorisation with elliptic curves as well as links to other web page.

>>>>

\newpage
\section{Elliptic Curves}
\index{Elliptic curves}
\hypertarget{ellcurve}{}
(Filipovics B. / B"uger M. / Esslinger B. / Oyono R., April 2000, Updates: Dec. 2001, June 2002, Mar. 2003)

@@

The secret key $s$ of the signature generator is a (random\index{Random}) whole number $s$ in
the interval $[1, r-1]$. The public key of the signature generator is a point
$W=(x,y)$ on the elliptic curve ${\bf E}$. The public key $W$ and secret key $s$
are interrelated as follows: $W = sG$. This means that the domain parameters
(particularly $G$) and the secret key $s$ are used to calculate the public key
$W$ (by adding $G$ $s$ times on ${\bf E}$). The ECDLP is obviously used here: If
$W$ and $G$ (as well as the other domain parameters used) are known, then it is
difficult to use these to calculate $s$. (If the parameters are chosen
correctly, this currently appears to be practically impossible).

@@

CrypTool\index{CrypTool} also offers elliptic curves for the digital signature function.

@@

%--------------------------------------------------------------------
\newpage
%\addcontentsline{toc}{subsection}{Literaturverzeichnis}
\begin{thebibliography}{99999}
\addcontentsline{toc}{subsection}{Bibliography}

\bibitem[Lenstra1987]{Lenstra1987} H.W. Lenstra\\ \index{Lenstra 1987}
   Factoring integers with elliptic curves, 
   Annals of Mathematics 126, pp. 649-673, 1987.

\end{thebibliography}

%--------------------------------------------------------------------
%\newpage
\section*{Web links}\addcontentsline{toc}{subsection}{Web links}

\begin{enumerate}
   \item Certicom Online Tutorial, \index{Certicom}\\
         \href{http://www.certicom.com/resources/ecc_tutorial/ecc_tutorial.html}{\texttt{http://www.certicom.com/resources/ecc\_tutorial/ecc\_tutorial.html}}

   \item IEEE P1363
          \href{http://grouper.ieee.org/groups/1363}{\texttt{http://grouper.ieee.org/groups/1363}}

   \item \hypertarget{Lenstra2}{}
         An informative web page about factorisation with elliptic curves. \\
         \href{http://www.loria.fr/~zimmerma/records/ecmnet.html}
	 {\texttt{http://www.loria.fr/\~{}zimmerma/records/ecmnet.html}} \\
         There, one finds literature to the topic factorisation with 
	 elliptic curves as well as links to other web page.

### 2003-03-10 script/en/chapters/introduction.tex
no message

% ............................................................................
%                         E I N F Ü H R U N G
% ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

\section*{Introduction}  \addcontentsline{toc}{section}{Introduction}

This script is delivered together with CrypTool\index{CrypTool}.

CrypTool is a program with an extremely comprehensive online help enabling you
to use and analyse cryptographic procedures within a unified graphical user interface.\par \vskip + 3pt

CrypTool was developed during the end-user awareness program in order to
increase employee awareness of IT security and provide them with a deeper
understanding of the term security.\par \vskip + 3pt

Since then CrypTool\index{CrypTool} has been used for education in companies and
universities and several universities help to develop further project.

The {\em authors} tried to describe cryptography for a broad audience --
without being mathematically incorrect. We believe, that this didactical
pretension is the best way to promote the awareness for IT security and the
readiness to use standardised modern cryptography.

Additionally we aimed to describe the newest results, as well as their
origins, concerning primes, factorization and cryptanalysis of AES, in order to be highly topical.
\\

At this point I'd like to thank explicitely 3 persons who espcially 
contributed to CrypTool. Without their talents and engagement 
CrypTool would not be what it is today:
\begin{itemize}
   \item Mr. Henrik Koy
   \item Mr. Joerg-Cornelius Schneider and
   \item Dr. Peer Wichmann.
\end{itemize}
Also I want to thank all the ones not named here for their engagement (mostly
performed in their spare time).
\\

@@

>>>>

% ............................................................................
%      V O R W O R T  und  E I N F Ü H R U N G (Zusammenspiel Skript-CT) 
% ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

% --------------------------------------------------------------------------
\section*{Preface to the 6th edition of the CrypTool script}  \addcontentsline{toc}{section}{Preface to the 6th edition of the CrypTool script}

Starting in the year 2000 this script became a part of the 
CrypTool\index{CrypTool} package. It should accomplish the program 
CrypTool by explaining some mathematical topics in more details, 
but still in a good to understand manner.

In order to also enable independent developers/authors work together 
the topics have been split up and for each topic an extra chapter was 
written readable by its own. The later editorial work in TeX added 
cross linkages and descriptions, where you can find the according 
functions within the CrypTool\index{CrypTool} program. Naturally 
there are much more interesting topics in mathematics and cryptography 
which could be more deepened -- therefore this is only one of many
ways to do it.

Rapid spread of the Internet also lead to an intensified research 
in the technologies involved. Especially within the area of cryptography 
much new knowledge came up.

This edition of the script updates the summaries of the following 
topical research areas:         
\begin{itemize}
\item the search for the largest prime numbers (generalized Mersenne 
      and Fermat primes), 
\item progress in number theory ("Primes in P"), 
\item the factorisation of big numbers (TWIRL) and 
\item progress in cryptanalysis of the AES standard. 
% \item usage of elliptic curves.
\end{itemize}

The first time when the script was delivered with CrypTool\index{CrypTool} 
was in version 1.2.01. Since then it was extended and updated with each 
new version of CrypTool (1.2.02, 1.3.00, 1.3.02, 1.3.03 and now 1.3.04).

I'd be more than happy if this keeps on going also in the further 
open source versions of CrypTool\index{CrypTool}.

I am deeply grateful to all the people helping with their impressive engagement to make this project successful and wide-spreaded.

I hope that many readers have fun with this script and that they get 
out of it more interest and higher understanding for this modern but 
also very ancient topic.

@@

% --------------------------------------------------------------------------
\newpage
\section*{Introduction -- How do the script and CrypTool play together?}  \addcontentsline{toc}{section}{Introduction -- How do the script and CrypTool play together?}

\textbf{This script}

This script is delivered together with CrypTool\index{CrypTool}. \par \vskip + 3pt

The {\em authors} tried to describe cryptography for a broad audience --
without being mathematically incorrect. We believe, that this didactical
pretension is the best way to promote the awareness for IT security and the
readiness to use standardised modern cryptography.
\par \vskip + 15pt

\textbf{The program CrypTool\index{CrypTool}}

CrypTool\index{CrypTool} is a program with an extremely comprehensive online help enabling you
to use and analyse cryptographic procedures within a unified graphical user interface.\par \vskip + 3pt

CrypTool\index{CrypTool} was developed during the end-user awareness program in order to
increase employee awareness of IT security and provide them with a deeper
understanding of the term security.\par \vskip + 3pt

Since then CrypTool\index{CrypTool} has often been used for 
education in companies and teaching and several universities 
help to further develop the project. \par \vskip + 3pt
\vskip +10 pt

At this point I'd like to thank explicitely 3 persons who especially 
contributed to CrypTool\index{CrypTool}. Without their talents and engagement 
CrypTool would not be what it is today:
\begin{itemize}
   \item Mr. Henrik Koy
   \item Mr. J"org-Cornelius Schneider and
   \item Dr. Peer Wichmann.
\end{itemize}
Also I want to thank all the many ones not named here for their 
engagement (mostly performed in their spare time).
\\
